---
title: "hw3"
author: "Piotr Kamoda"
date: "22 pazdziernika 2016"
output: html_document
---


```{r setup, echo=FALSE}
library(rpart)
library(randomForest)
d1=read.table("student-mat.csv",sep=";",header=TRUE)
d1Train <- d1[0:300,]
d1Test <- d1[301:600,]
d1Train$Dalc <- factor(d1Train$Dalc) #Make sure setup is ok
d1Test$Dalc <- factor(d1Test$Dalc) #Make sure setup is ok
d1$Dalc <- factor(d1$Dalc)

```

The task is to compare random forest and decision trees on the basis of their performance. Let's start with decision trees:

```{r dt}
rtree <- rpart(Dalc ~ school+sex+age+famsize+reason+failures+romantic, data = d1Train, method="class")
tab <- table(real = d1Test$Dalc,
      predicted = predict(rtree, data=d1Test, type = "class"))
```

Lets look at the output:

```{r dto, echo=FALSE}
show(rtree)
show(tab)
plot(rtree)
text(rtree)
```

Let's check quality of the prediction: 
```{r dtq, echo=FALSE}
sum(diag(tab)) / sum(tab)
```

Quality of this prediction is 0.6. That's not a very good prediction.

Now we can go on to random forests:

```{r rf}
ffit <- randomForest(Dalc ~ school+sex+age+famsize+reason+failures+romantic,   data=d1Train, importance = TRUE)
print(ffit) 
```

Lets look at some details of how this model looks like:

```{r rfd, echo=FALSE}
importance(ffit)
varImpPlot(ffit)
head(predict(ffit, type="prob"))
```

And check the quality:

```{r rfq, echo=FALSE}
sum(diag(ffit$confusion)) / sum(ffit$confusion)
```

Quality of the prediction exceeded 0.7. <br />

Just to make sure let's check models with less variables:

```{r both}
rtree <- rpart(Dalc ~ failures+absences+freetime, data = d1Train, method="class")
ffit <- randomForest(Dalc ~ failures+absences+freetime,   data=d1Train, importance = TRUE)
```

And their qualities (in order):

```{r bothq, echo=FALSE}
tab <- table(real = d1Test$Dalc,
      predicted = predict(rtree, data=d1Test, type = "class"))
sum(diag(tab)) / sum(tab)
sum(diag(ffit$confusion)) / sum(ffit$confusion)
```

Smaller difference but still visible.

### Conclusion
Random forests have better quality of final prediction then decision trees.